---
title: "Dataquest Project: Exploratory Visualization For Forest Fire Data"
author: Nneamaka Abigail Chalokwu
date: 9th july 2020
output: html_notebook
---

```{r}

```
Load the necessary libraries
```{r load libraries}
# load packages
library(readr)
library(dplyr)
library(ggplot2)
library(purrr)
```


```{r}
```
import the dataset
```{r import data and save as a dataframe}
forest_fire_data <- read.csv("forestfires.csv")
```
Group the data by month and days of the week
```{r}
#group the data by months
forest_data_month <- forest_fire_data %>%
  group_by(month) %>%
  summarize(total_no_fires = n())
#visualize using barchart
ggplot(data = forest_data_month,
       aes(x = month, y =total_no_fires)) + geom_bar(stat = "identity") + 
       theme(panel.background = element_rect(fill = "azure")) + labs(title = "Monthly Record of Forest Fire")
```
```{r}
#groupby days of the week
forest_data_days <- forest_fire_data %>% 
  group_by(day) %>%
  summarise(total_no_fire = n())
#visualise using barchart
ggplot(data = forest_data_days,
       aes(x = day, y = total_no_fire)) + 
       geom_bar(stat = "identity") +
       theme(panel.background = element_rect(fill = "cornsilk")) + labs(title = "Daily Record Of Forest Fire")

```
```{r}
```

rearrange the months and days in a chronological order
```{r}
# rearrange the months and day column in a chronological order
forest_fire_data <- forest_fire_data %>%
  mutate(month = factor(month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")),
         day = factor(day, levels = c("sun", "mon", "tue", "wed", "thu", "fri", "sat")))

```
Regroup and visualize the rearranged data
```{r group and visualize the reodered data}
forest_data_month <- forest_fire_data %>%
  group_by(month) %>%
  summarize(total_no_fires = n())

ggplot(data = forest_data_month,
       aes(x = month, y =total_no_fires)) + geom_bar(stat = "identity") + 
       theme(panel.background = element_rect(fill = "thistle")) + labs(title = "Monthly Record of Forest Fire")
```
```{r}
forest_data_days <- forest_fire_data %>% 
  group_by(day) %>%
  summarise(total_no_fire = n())

ggplot(data = forest_data_days,
       aes(x = day, y = total_no_fire)) + 
       geom_bar(stat = "identity") +
       theme(panel.background = element_rect(fill = "coral")) + labs(title = "Daily Record Of Forest Fire")


```
create a function to visualize using boxplots
```{r}
box_plots <- function(x,y) {
  ggplot(data = forest_fire_data) +
  aes_string(x = x, y = y) +
    geom_boxplot() +
    theme(panel.background = element_rect(fill = "white"))
}
```
Assign variable names to X and Y
```{r}
x_month_var <- names(forest_fire_data)[3]
x_day_var <- names(forest_fire_data)[4]
y_var <- names(forest_fire_data)[5:12]
```
Apply map() function to the variables created above
```{r}
Box_for_month <- map2(x_month_var, y_var,box_plots)
Box_for_day <- map2(x_day_var, y_var,box_plots)
```
create a function for scatter plots
```{r}
scatter_plots <- function(x,y) {
  ggplot(data = forest_fire_data) +
    aes_string(x = x, y = y) +
    geom_point() +
    theme(panel.background = element_rect(fill = "white"))
}
```
Assign variables to X and Y
```{r}
x_var <- names(forest_fire_data)[5:12]
y_var <- names(forest_fire_data)[13]

```
Apply Map() to the variables
```{r}
the_scatters <- map2(x_var, y_var, scatter_plots)
```

